\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-11-13 5-2-aus-hist-max-temp.Rmd
more updates

```{r echo = TRUE}
tmax <- tmax |>
  face_temporal() |> 
  group_by(month = lubridate::month(date),
         group = as.factor(
           ifelse(lubridate::year(date) > 2015,
           "2016 ~ 2020", "1971 ~ 1975"))) |>
  summarise(tmax = mean(tmax, na.rm = TRUE))
```

@@

p1 <- ggplot() + 
  geom_sf(data = nsw_vic, fill = "grey95", 
          color = "white", size=2) + 
  geom_glyph(
    data = tmax, 
    aes(x_major = long, x_minor = month, 
        y_major = lat, y_minor = tmax,
        group = interaction(id, group), color = group),
    width = 1.2, height = 0.8) +
    geom_rect(
      data = box_df,
      aes(xmin = long - 0.6, xmax = long + 0.6,
          ymin = lat - 0.12, ymax = lat + 0.35),
      fill = "transparent", color = "black") +
  scale_color_brewer("", palette = "Dark2") + 
  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
  ggthemes::theme_map() +
  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
  guides(color = guide_legend(override.aes = list(size=2)))

>>>>

```{r echo = TRUE}
tmax <- tmax |>
  face_temporal() |> 
  mutate(month = lubridate::month(date),
         group = as.factor(
           ifelse(lubridate::year(date) > 2015,
           "2016 ~ 2020", "1971 ~ 1975"))) |>
  group_by(month, group) |> 
  summarise(tmax = mean(tmax, na.rm = TRUE))
```

@@

p1 <- ggplot() + 
  geom_sf(data = nsw_vic, fill = "grey95", 
          color = "white") + 
  geom_glyph(
    data = tmax, 
    aes(x_major = long, x_minor = month, 
        y_major = lat, y_minor = tmax,
        group = interaction(id, group), color = group),
    width = 1.2, height = 0.8) +
    geom_rect(
      data = box_df,
      aes(xmin = long - 0.6, xmax = long + 0.6,
          ymin = lat - 0.12, ymax = lat + 0.35),
      fill = "transparent", color = "black") +
  scale_color_brewer("", palette = "Dark2") + 
  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
  ggthemes::theme_map() +
  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
  guides(color = guide_legend(override.aes = list(size=2)))

### 2022-11-13 jss/main.tex
more updates

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax |>
+   face_temporal() |> 
+   group_by(month = lubridate::month(date),
+          group = as.factor(
+            ifelse(lubridate::year(date) > 2015,
+            "2016 ~ 2020", "1971 ~ 1975"))) |>
+   summarise(tmax = mean(tmax, na.rm = TRUE))
\end{CodeInput}
\end{CodeChunk}

@@

The ERA5 data \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF). It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \pkg{ecmwfr} package \citep{ecwmfr}. For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the \emph{specific humidity} and \emph{geopotential}. The downloaded NetCDF data (\texttt{raw}) is read into \proglang{R} using the \pkg{ncdf4} package, and converted to a \code{cubble} object:

>>>>

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax |>
+   face_temporal() |> 
+   mutate(month = lubridate::month(date),
+          group = as.factor(
+            ifelse(lubridate::year(date) > 2015,
+            "2016 ~ 2020", "1971 ~ 1975"))) |>
+   group_by(month, group) |> 
+   summarise(tmax = mean(tmax, na.rm = TRUE))
\end{CodeInput}
\end{CodeChunk}

@@

The ERA5 data \citep{hersbach2020era5} provides hourly estimates across the Earth for atmospheric, land and oceanic climate variables. The data is available in the NetCDF format from the European Centre for Medium-Range Weather Forecasts (ECMWF). It can be directly downloaded from Copernicus Climate Data Store (CDS) \citep{cds} website or via the \pkg{ecmwfr} package \citep{ecwmfr}. For the reproduction, we focus on the \code{era5-pressure} data, hourly pressure levels from 1970 to present, with the \emph{specific humidity} and \emph{geopotential}. The downloaded NetCDF data (\texttt{raw}) is first converted to a \code{cubble} object:
